### 2013-11-12 src2/thesis.tex
spell pass

\noindent The {\tt vecAdd} program takes two {\tt SPull} input arrays and gives one 
as result. The Obsidian {\tt vecAdd} is more general than the CUDA version. From it 
we can generate CUDA kernels for any element type that supports the {\tt (+)} operation. 
In Haskell we know that types that belong to the {\tt Num} typeclass has a {\tt (+)} operation.

@@

We have implemented basic algorithms such as sorting, merging, fractal generation, reductions 
and parallel prefix (scans). Often we have managed to get very good performance, close to that 
of hand optimised CUDA code. GPU execution of scan operations is something that has 
been worked on extensively in the GPGPU community and very efficient implementations 
exist~\citet{merrill,ScanCUDA,OLAMARCUS,EFFICIENTSCAN}. In Paper F (section~\ref{sec:paperF}), 
we only study scan kernels (scan of small arrays) and measure their speed relative to each other. 
Compared to a scan kernel supplied with the CUDA API, our fastest kernel is slighty more 
than three times slower. Performance needs to be improved at the kernel level before we attempt 
composing them into large scan operations. We hope to improve this situation with the 
addition of the warp level abstractions in Obsidian (this is future work, 
section~\ref{sec:thesisFUTURE}).

Paper F (section~\ref{sec:paperF}) tells a story about how to implement and tune 
a reduction kernel in Obsidian. Seven reductions kernels are implemented, the first 
very simple and naive. Each following kernel, tweaks some aspect of the previous one. 
Applying these changes is pleasingly simple using Obsidian; some of these changes 
would mean a major rewrite in CUDA and some are even impossible. 
The final kernel reaches a throughput of 140GB/sec on a GPU with 192GB/sec memory bandwith. 
At this point, we have pushed Obsidian as far as possible with the current abstractions. 
However, it is possible that with more work on warp level capabilites, this could be 
improved a bit further.   
%faster than the reduction skeleton used in Accelerate. This is quite a feat, since 
%reduction in Accelerate corresponds to just one single hand optimised skeleton. We would have 
%been happy just getting close.

@@

>>>>

\noindent The {\tt vecAdd} program takes two {\tt SPull} input arrays and gives one 
as result. The Obsidian {\tt vecAdd} is more general than the CUDA version. From it 
we can generate CUDA kernels for any element type that supports the {\tt (+)} operation. 
In Haskell we know that types that belong to the {\tt Num} type class has a {\tt (+)} operation.

@@

We have implemented basic algorithms such as sorting, merging, fractal generation, reductions 
and parallel prefix (scans). Often we have managed to get very good performance, close to that 
of hand optimised CUDA code. GPU execution of scan operations is something that has 
been worked on extensively in the GPGPU community and very efficient implementations 
exist~\citet{merrill,ScanCUDA,OLAMARCUS,EFFICIENTSCAN}. In Paper F (section~\ref{sec:paperF}), 
we only study scan kernels (scan of small arrays) and measure their speed relative to each other. 
Compared to a scan kernel supplied with the CUDA API, our fastest kernel is slightly more 
than three times slower. Performance needs to be improved at the kernel level before we attempt 
composing them into large scan operations. We hope to improve this situation with the 
addition of the warp level abstractions in Obsidian (this is future work, 
section~\ref{sec:thesisFUTURE}).

Paper F (section~\ref{sec:paperF}) tells a story about how to implement and tune 
a reduction kernel in Obsidian. Seven reduction kernels are implemented, the first 
very simple and naive. Each following kernel tweaks some aspect of the previous one. 
Applying these changes is pleasingly simple using Obsidian; some of these changes 
would mean a major rewrite in CUDA. The final kernel is almost 5 times as fast as the first, 
and reaches a throughput of 140GB/sec on a GPU with 192GB/sec memory bandwidth. The speed of any 
kernel will always be limited by how fast the data can be pushed through the GPU; this performance 
is quite satisfactory. In Paper F, a large reduction algorithm based on this kernel is 
implemented that runs slightly faster than Accelerates fold skeleton. With this reduction 
case study, we have pushed Obsidian as far as possible with the current abstractions. 
However, it is possible that with more work on the warp level abstractions, this could be 
improved further.   
%faster than the reduction skeleton used in Accelerate. This is quite a feat, since 
%reduction in Accelerate corresponds to just one single hand optimised skeleton. We would have 
%been happy just getting close.

@@

An in depth comparison of Obsidian and Accelerate is hard to perform. The reason for 
this is that Obsidian is a language for kernels, while Accelerate is a language for 
applications. Any larger GPU application consists of a number of kernels that 
are cooperating. Obsidian only has very basic support for launching and coordinating 
kernels for such applications, from within Haskell and Accelerate is limited in specifying 
new specialised kernels.
